---
title: "Using gumboot"
author: "Martyn Clark and Kevin Shook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(gumboot)
```

Clark et al. (2021) have shown how Goodness of Fit (GOF) statistics for
hydrological models can easily be misused. The purpose of `gumboot` is to
evaluate the error in GOF statistics, through jacknifing
after bootstrapping.

## Theory
I'm leaving this for you, Martyn!

## How to use `gumboot`
The most important function is `bootjack()` which computes the bootstap and jackknife statistics
for a single data set, which is a data frame containing dates, observed
and simulated flows.

A data set (`flows_1030500`) is provided.
```{r}
flows_1030500 <- flows_1030500 
head(flows_1030500)
```

Plotting the values shows fairly good agreement between the simulated and observed values  

```{r, fig.width = 5}
library(ggplot2)
library(reshape2)
melted <- melt(flows_1030500, id.vars = "date")
ggplot(melted, aes(date, value, colour = variable)) +
  geom_line() +
  xlab("") +
  ylab("Daily streamflow (m³/s)") 
```

To perform the bootstrap and jackknife analyses, the values are 
passed to `bootjack`. There are many options.
The default is to calculate statistics for both NSE (Nash-Sutcliffe efficiency), and
KGE (Kling-Gupta efficiency) values. In this example, we will compute the
statistics of the NSE.  

Note that the name of the observed variable must be `obs` and the name
of the simulated variable must be `sim`.

```{r}
NSE_values <- bootjack(flows_1030500, GOF_stat = "NSE")
NSE_values
```
In this example, the standard error of the NSE statistic, as calculated by jackknifing is `r NSE_values$seJack[1]`; as calculated by bootstrapping it is `r NSE_values$seBoot[1]`, and as calculated
by jackknifing after the bootstrapping (JAB) is `r NSE_values$seJab[1]`, showing the uncertainty
in the statistic.  

It is important to note that the bootstrap and jackknife-after-bootstrap standard errors are dependant
on random samples of the years, so the values will change with each execution of the function.
If you want to get the same values each time, for example to compare your results with
another person's analyses, you have two options.

If you set the option `seed` to have a value, the **R** random number generator will always
return the same sequence of values, as shown below:  

```{r}
bootjack(flows_1030500, GOF_stat = "NSE", seed = 1)
bootjack(flows_1030500, GOF_stat = "NSE", seed = 1)
```

Note that the value of `seJack` above is identical to the previously determined value,
as the jackknifing always uses the same set of years of data.


If you want to compare the results with code not written in **R** you can save the randomly selected
years to a file. If the specified file does not exist, it will be written to. If
the file does exist, then the years will be read from it. 

### Raw values
If you are interested in the values used to calculate the standard errors, you can return them
using the option `returnSamples = TRUE`. The function returns a list with the values
for the bootstrap and jackknifing analyses.

```{r}
NSE_samples <- bootjack(flows_1030500, GOF_stat = "NSE", returnSamples = TRUE)
names(NSE_samples)
```

You can see the variability of the NSE values (as well as the sampled observed and 
simulated values) as determined by the bootstrap and jackknife.

```{r, fig.width=5, warning=FALSE, message = FALSE, error=FALSE}
ggplot(NSE_samples$statsBoot, aes(NSE)) + 
  geom_histogram() +
  ggtitle("Bootstrap samples")
```


## Multiple locations

The function `conus_hcdn_bootjack()` applies `bootjack()` to model runs over the
continental US (CONUS). The model runs are not supplied, but would need to
be stored in a NetCDF file.

Note - I think we'll need a description of the HCDN dataset here.

